---
title: "p8105_hw6_kpp2126"
author: "Kevin P. Patterson"
date: "2022-11-26"
output: github_document
---

```{r libraries}
library(tidyverse)
```

## Problem 1

```{r weather data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. 
1. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:
*r̂ 2
*log(β̂ 0∗β̂ 1)

* Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities.
```{r 5k bootstrap samples}

```

2. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1)
*Note: `broom::glance()` is helpful for extracting r̂ 2 from a fitted regression, and`broom::tidy()` (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).




